---
title: "Analysis with Conos"
author: "jeprob"
date: "2021-01-07"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Conos integration: 
Dataset preprocessing and integration was adapted from [this vignette](http://htmlpreview.github.io/?https://github.com/satijalab/seurat-wrappers/blob/master/docs/conos.html) on the Seurat github. The Conos integration followed [this vignette](https://github.com/kharchenkolab/conos/blob/master/vignettes/walkthrough.md) by the developers of Conos, subsection: 'Systematic comparative analysis of human PBMC'. 


## Load packages
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(conos)
library(dplyr)
library(pagoda2)
library(Seurat) 
```


## Read in the data and set up Seurat object
```{r echo=TRUE, message=FALSE, warning=FALSE}
data_pbmc <- readRDS("./data/wonglab_10x_XLP2_sce_after_scDblFinder.rds")
data_pbmc
head(counts(data_pbmc))
```


## Set up a Seurat object and preprocess with Seurat
```{r echo=TRUE, message=FALSE, warning=FALSE}
#set up seurat object
pbmc <- as.Seurat(x=data_pbmc)
pbmc
#split into list of individual datasets
pbmc.panel <- SplitObject(pbmc, split.by = 'Sample')
str(pbmc.panel,1)
```


## Preprocessing with Seurat
Normal Seurat preprocessing steps are NormalizeDate, FindVariableFeatures, ScaleData and run PCA and TSNE for each dataset. 
```{r echo=TRUE, message=FALSE, warning=FALSE}
for (i in 1:length(pbmc.panel)){
  pbmc.panel[[i]] <- NormalizeData(pbmc.panel[[i]]) %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA(verbose=FALSE) %>% RunTSNE(verbose=FALSE)
}
```

## Integrating dataset with Conos
```{r echo=TRUE, message=FALSE, warning=FALSE}
# construct Conos object for the panel of samples
pbmc.con <- Conos$new(pbmc.panel)
pbmc.con #have a look at the Conos object
str(pbmc.con$samples,1)
```


We now build the joint graph: 
```{r echo=TRUE, message=FALSE, warning=FALSE}
pbmc.con$buildGraph(k=15, k.self=5, space='PCA', ncomps=30, n.odgenes=2000, matching.method='mNN', metric='angular', score.component.variance=TRUE, verbose=TRUE)
```

Next, we can visualize the estimated amount of variance explained by successive PCs:
```{r echo=TRUE, message=FALSE, warning=FALSE}
plotComponentVariance(pbmc.con, space='PCA')
```
We can see that the first PC already explaines a large fraction of the variance. The second and the third PC also explain a much higher fraction of variance than the remaining ones. 


### Downstream analysis
We can now use the graph to find and visualize global clusters. Here, we use the Leiden community detection.
```{r echo=TRUE, message=FALSE, warning=FALSE}
pbmc.con$findCommunities(method=leiden.community, resolution=0.8)
```

As a next step, the joint graph can be used to visualize the integration of datasets and the clustering.
```{r}
pbmc.con$plotGraph(color.by='sample', mark.groups=FALSE, alpha=0.1, show.legend=TRUE)
```
One can observe that the data is integrated in a joint TSNE space and cells do not cluster by dataset. 

```{r}
pbmc.con$plotGraph(alpha=0.5)
```
We find that there are 14 clusters identified in total. 

Next, the joint clustering for each panel are plotted, where cell labels obtained from the joint clustering are added in the original TSNE embedding. 
```{r}
pbmc.con$plotPanel(font.size=4)
```
We find different amounts of cell types in the individual datasets. To have a closer look at the composition of the individual datasets, we plot cluster barplots for the individual datasets. 
```{r}
plotClusterBarplots(pbmc.con, legend.height = 0.1)
```
Many interesting things can be observed in this plot. It is noteable that most of the cells in ko1 belong to cluster 2. Furthermore, the distribution of cells to clusters differs from wt1 to wt2, judged by eye. Cluster 12 is not present in the dataset het1.

Further downstream analysis such as gene expression visualization or hierarchical community structure generation would be possible with Conos. For more information visit [this vignette](https://github.com/kharchenkolab/conos/blob/master/vignettes/walkthrough.md).