---
title: "Analysis with Conos"
author: "jeprob"
date: "2021-01-07"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Conos integration: 
Dataset preprocessing and integration was adapted from [this vignette](http://htmlpreview.github.io/?https://github.com/satijalab/seurat-wrappers/blob/master/docs/conos.html) on the Seurat github, subsection: 'Systematic comparative analysis of human PBMC'. The Conos integration followed [this vignette](https://github.com/kharchenkolab/conos/blob/master/vignettes/walkthrough.md) by the developers of Conos. 


## Load packages
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(conos)
library(dplyr)
library(pagoda2)
library(Seurat) 
set.seed(333)
```


## Data readin
```{r echo=TRUE, message=FALSE, warning=FALSE}
data_pbmc <- readRDS("./data/wonglab_10x_XLP2_sce_after_scDblFinder.rds")
data_pbmc
head(counts(data_pbmc))
```


## Set up a Seurat object 
```{r echo=TRUE, message=FALSE, warning=FALSE}
#set up seurat object
pbmc <- as.Seurat(x=data_pbmc)
pbmc
#split into list of individual datasets
pbmc.panel <- SplitObject(pbmc, split.by = 'Sample')
str(pbmc.panel,1)
```


## Preprocessing with Seurat
Normal Seurat preprocessing steps are NormalizeDate, FindVariableFeatures, ScaleData and run PCA and t-SNE for each dataset. 
```{r echo=TRUE, message=FALSE, warning=FALSE}
for (i in 1:length(pbmc.panel)){
  pbmc.panel[[i]] <- NormalizeData(pbmc.panel[[i]]) %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA(verbose=FALSE) %>% RunTSNE(verbose=FALSE)
}
```

## Integration with Conos
```{r echo=TRUE, message=FALSE, warning=FALSE}
# construct Conos object for the panel of samples
pbmc.con <- Conos$new(pbmc.panel)
pbmc.con #have a look at the Conos object
str(pbmc.con$samples,1)
```


We now build the joint graph: 
```{r echo=TRUE, message=FALSE, warning=FALSE}
pbmc.con$buildGraph(k=15, k.self=5, space='PCA', ncomps=30, n.odgenes=2000, matching.method='mNN', metric='angular', score.component.variance=TRUE, verbose=TRUE)
```

Next, we can visualize the estimated amount of variance explained by successive PCs:
```{r echo=TRUE, message=FALSE, warning=FALSE}
plotComponentVariance(pbmc.con, space='PCA')
```
We can see that the first PC already explaines a large fraction of the variance. The second and the third PC also explain a much higher fraction of variance than the remaining ones. 


## Downstream analysis
We can now use the graph to find and visualize global clusters. Here, we use the Leiden community detection and visulatization in a two-dimensional space using the largeVis algorithm, which is default for Conos.
```{r echo=TRUE, message=FALSE, warning=FALSE}
pbmc.con$findCommunities(method=leiden.community, resolution=0.8)
```

As a next step, the joint graph can be used to visualize the integration of datasets and the clustering.
```{r}
pbmc.con$plotGraph(color.by='sample', mark.groups=FALSE, alpha=0.1, show.legend=TRUE)
```
One can observe that the data is well integrated in the space and cells do not cluster by dataset. 

```{r}
pbmc.con$plotGraph(alpha=0.5)
```
We find that there are 13 clusters identified in total. 

Next, cells in the original t-SNE embedding were plotted for each panel and cell labels obtained from the joint clustering were added.
```{r}
pbmc.con$plotPanel(font.size=4)
```
We observe that different amounts of the individual cell types are present in the different samples. To have a closer look at the composition of the individual samples, we plot cluster barplots for the individual datasets. 
```{r}
plotClusterBarplots(pbmc.con, legend.height = 0.1)
```
Many interesting things can be observed in this plot. The number of cells belonging to the clusters has a big range from many 1000 to roughly 50 cells. Cells from smaller clusters still seemed to be observed in all samples, except for cluster 12 which does not appear in het1. It is notable that most of the cells in ko1 belong either to cluster 2. Furthermore, there is no obvious trend, judged by eye, that wildtype sample cells seemed to be differently distributed than the remaining samples het1, m2 and p2. 

Further downstream analysis such as gene expression visualization or hierarchical community structure generation would be possible with Conos. For more information visit [this vignette](https://github.com/kharchenkolab/conos/blob/master/vignettes/walkthrough.md).
